---
title: "Batch processing with polars"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Batch processing with polars}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

For large sequence datasets, immunum provides a vectorized
[polars](https://pola.rs/) integration that runs numbering and
segmentation in parallel across all available cores using Rust's rayon
thread pool. This mirrors the Python `immunum.polars` module one-to-one.

## Setup

Install polars from r-multiverse:

```{r, eval = FALSE}
install.packages("polars",
  repos = c("https://community.r-multiverse.org", getOption("repos"))
)
```

## Numbering

`polars_number()` returns a polars expression that produces a struct
column with fields `chain`, `scheme`, `confidence`, `positions`, and
`residues`:

```{r}
library(immunum)
library(polars)

df <- pl$DataFrame(sequence = c(
  "QVQLVQSGAEVKRPGSSVTVSCKASGGSFSTYALSWVRQAPGRGLEWMGGVIPLLTITNYAPRFQGRITITADRSTSTAYLELNSLRPEDTAVYYCAREGTTGKPIGAFAHWGQGTLVTVSS",
  "DIQMTQSPSSLSASVGDRVTITCRASQDVNTAVAWYQQKPGKAPKLLIYSASFLYSGVPSRFSGSRSGTDFTLTISSLQPEDFATYYCQQHYTTPPTFGQGTKVEIK"
))

result <- df$with_columns(
  polars_number("sequence", chains = c("H", "K", "L"), scheme = "imgt")$
    alias("numbered")
)

# Unnest the struct to get individual columns
result$unnest("numbered")
```

The first argument can be a polars expression (`pl$col("sequence")`)
or a column name string — the string shortcut is equivalent to
`pl$col(...)`.

### Accessing fields

```{r}
# Chain detection result for every row
result$get_column("numbered")$struct$field("chain")

# Position lists (List[String])
result$get_column("numbered")$struct$field("positions")
```

## Segmentation

`polars_segment()` returns a struct with the nine FR/CDR region strings:

```{r}
seg <- df$with_columns(
  polars_segment("sequence", chains = c("H", "K", "L"), scheme = "imgt")$
    alias("segmented")
)

# Unnest to separate columns
seg$unnest("segmented")
# columns: prefix, fr1, cdr1, fr2, cdr2, fr3, cdr3, fr4, postfix
```

## Reusing an Annotator

When you need to apply the same numbering configuration across multiple
queries or expressions, build the `Annotator` once and pass it to the
`*_method()` variants. This skips the per-batch annotator construction
step:

```{r}
ann <- Annotator$new(chains = c("H", "K", "L"), scheme = "imgt")

# Numbering
df$with_columns(
  polars_numbering_method("sequence", annotator = ann)$alias("numbered")
)

# Segmentation
df$with_columns(
  polars_segmentation_method("sequence", annotator = ann)$alias("segmented")
)
```

This mirrors the Python `immunum.polars.numbering_method()` /
`immunum.polars.segmentation_method()` API.

## Performance notes

- **Thread parallelism** is handled by rayon on the Rust side.
  Each batch fans out across all available cores automatically.
- **Batch size**: polars passes the entire column as a single batch
  to `map_batches`. There is one R↔C boundary crossing per batch,
  not per row, so the overhead is negligible.
- **Memory**: numbering results are columnar (separate `positions`
  and `residues` List columns) rather than a Map, matching the
  upstream Python struct layout.
- **Failed sequences** (NA, alignment failure, sub-threshold
  confidence) produce a row of polars nulls. Row count is always
  preserved.

## Minimum confidence

All four functions accept an optional `min_confidence` parameter
(numeric in `[0, 1]`). Sequences below this threshold are returned as
null rows. The default (`NULL`) uses the Rust core's default of 0.5:
```{r}
# Accept all alignments regardless of confidence
df$with_columns(
  polars_number("sequence",
                chains = c("H", "K", "L"),
                scheme = "imgt",
                min_confidence = 0.0)$alias("numbered")
)
```
