---
title: "Getting started with immunum"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with immunum}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

`immunum` is an R package for numbering antibody and T-cell receptor (TCR)
variable-domain sequences using IMGT and Kabat schemes. It wraps the
[immunum Rust crate](https://crates.io/crates/immunum) via
[extendr](https://extendr.github.io/) for high performance.

## Installation

Install from r-universe:

```{r, eval = FALSE}
install.packages("immunum",
  repos = c("https://enpicom.r-universe.dev", getOption("repos"))
)
```

## Numbering

Create an `Annotator` and number a single sequence:

```{r}
library(immunum)

ann <- Annotator$new(chains = c("H", "K", "L"), scheme = "imgt")

sequence <- paste0(
  "QVQLVQSGAEVKRPGSSVTVSCKASGGSFSTYALSWVRQAPGRGLEWMGG",
  "VIPLLTITNYAPRFQGRITITADRSTSTAYLELNSLRPEDTAVYYCAREGT",
  "TGKPIGAFAHWGQGTLVTVSS"
)

result <- ann$number(sequence)
result$chain       # "H"
result$confidence  # 0.78
result$numbering   # named character: "1"="Q", "2"="V", ...
```

The `chains` argument controls which chain types are tried during alignment.
Chain aliases are case-insensitive — `"H"`, `"heavy"`, and `"IGH"` all
resolve to the same chain.

### Supported chains

| Antibody      | TCR          |
|---------------|--------------|
| IGH (heavy)   | TRA (alpha)  |
| IGK (kappa)   | TRB (beta)   |
| IGL (lambda)  | TRD (delta)  |
|               | TRG (gamma)  |

### Numbering schemes

- **IMGT** — all 7 chain types
- **Kabat** — antibody chains only (IGH, IGK, IGL)

## Segmentation

`segment()` splits the sequence into framework (FR) and
complementarity-determining (CDR) regions:

```{r}
seg <- ann$segment(sequence)
seg$fr1      # "QVQLVQSGAEVKRPGSSVTVSCKAS"
seg$cdr1     # "GGSFSTYA"
seg$fr2      # "LSWVRQAPGRGLEWMGG"
seg$cdr2     # "VIPLLTIT"
seg$fr3      # "NYAPRFQGRITITADRSTSTAYLELNSLRPEDTAVYYC"
seg$cdr3     # "AREGTTGKPIGAFAH"
seg$fr4      # "WGQGTLVTVSS"
```

Prefix and postfix residues that fall outside the numbered domain are
returned in `seg$prefix` and `seg$postfix`.

## Validation

immunum provides access to benchmark accuracy thresholds from the
upstream test suite:

```{r}
# List all known validation fixtures
validation_fixtures()

# Look up the benchmark-suite accuracy threshold
benchmark_threshold("imgt.H")   # 99.88
```

The validation fixture CSVs (reference numberings for ~10k sequences)
live at `fixtures/validation/` in the source repo and are available
when running tests from the repo tree -- same as the Python package.
